Predicting computational reproducibility of data analysis pipelines \\in large population studies using collaborative filtering

### 2018-07-30 paper.tex
Added effect of number of factors

\todo{Soudabeh, can you remove the caption on all matrices?}

@@

Two experiments have been conducted for each reproducibility matrix, to evaluate the performance of our 
predictions using (1) ALS without biases, and (2) ALS with biases. 
Results will be evaluated using classification accuracy and compared to (1) a dummy classifier that always 
predicts the value in the majority class and (2) the Random Unreal 
method, used as the baseline sampling technique. All reported values are averages
over 5 repetitions. Due to sampling issues, it is possible that the actual 
training ratio obtained with some of the sampling methods does not 
match the target one. We checked that the difference between the target 
and real training ratios was lower than 0.01. We used Spark's ALS model 
as available in package \texttt{pyspark.ml}, with 50 factors, a 
regularization parameter of 0.01, a maximum of 5 iterations and non 
negative constraints set to true. \todo{Tristan: Add one plot with sensitivity and specificity for 0.9 training}

@@

Accuracy results for the different numbers of subject types are 
reported in Figure~\ref{fig:results-synthetic}. By construction, the 
accuracy of the dummy classifier is close to 0.5 for all subject types \todo{Soudabeh: why such variations?}. Random 
Unreal performs very well for all subject types, which confirms that 
ALS is working correctly. Unsurprisingly, all the other methods perform 
better than the dummy classifier, and their accuracy decreases as the 
number of subject types increases.

@@

\subsubsection{ROC analysis}

@@

\subsection{Effect of the number of factors}
\todo{Soudabeh: Different number of factors: type 8, RFNU, 50, 100, 200, 400.}

@@

Figure~\ref{fig:overlap} illustrates the comparison between RFNU, RFNTL 
and RFNTS. The matrices in this Figure represent the union between the 
training and the test set, where the training set is represented in 
black (negatives) and white (positives), and the test set is 
represented in green (true positives), yellow (false negatives), gray 
(true negatives) and red (false positives). This representation 
provides insights regarding where, and perhaps why, prediction errors 
occur. At this training ratio, RFNU (Figure~\ref{fig:overlap-rfntu}) 
uniformly samples the number of files per subject between 0.8$N_f$ and 
$N_f$. It provides an opportunity for ALS to be trained on the last 
files of the pipelines, while maintaining a low number of columns with 
a low training ratio. On the contrary, RFNTL 
(Figure~\ref{fig:overlap-rfntl}) samples the number of files per 
subject from 0.85$N_f$, but the probability to have a complete column 
is 0 (the one complete column in Figure~\ref{fig:overlap-rfntl} is the 
one included to prevent cold start issues), which leads to a "stripe" 
of errors at the bottom of the matrix. RFNTS 
(Figure~\ref{fig:overlap-rfnts}) doesn't have this issue, as it 
includes many complete columns in the training set, but this comes at 
the cost of several columns with a very low number of files, lower than 
60\%. For such columns, the prediction may go somewhat wrong, which explains
the reduced accuracy compared to RFNU.

>>>>

@@

Two experiments have been conducted for each reproducibility matrix, to evaluate the performance of our 
predictions using (1) ALS without biases, and (2) ALS with biases. 
Results will be evaluated using classification accuracy and compared to (1) a dummy classifier that always 
predicts the value in the majority class and (2) the Random Unreal 
method, used as the baseline sampling technique. All reported values are averages
over 5 repetitions. Due to sampling issues, it is possible that the actual 
training ratio obtained with some of the sampling methods does not 
match the target one. We checked that the difference between the target 
and real training ratios was lower than 0.01. We used Spark's ALS model 
as available in package \texttt{pyspark.ml}, with 50 factors, a 
regularization parameter of 0.01, a maximum of 5 iterations and non 
negative constraints set to true.

@@

Accuracy results for the different numbers of subject types are 
reported in Figure~\ref{fig:results-synthetic}. By construction, the 
accuracy of the dummy classifier is close to 0.5 for all subject types. 
Random Unreal performs very well for all subject types, which confirms 
that ALS is working correctly. Unsurprisingly, all the other methods 
perform better than the dummy classifier, and their accuracy decreases 
as the number of subject types increases.

@@

\subsection{ROC analysis}

@@

\subsection{Varying the number of factors}
Figure~\ref{fig:factors} shows the effect of the number of factors used 
in ALS, for the RFNU method and the synthetic dataset. We do not find a 
clear effect of the number of factors on the accuracy.

\begin{figure}
\includegraphics[width=\columnwidth]{data/results/means_of_results/ALS/Synthetic/8type-RFNU-Differ-Num-Factors/ALS-8-types-ranks.pdf}
\caption{Effect of the number of factors used in ALS, RFNU method, synthetic dataset.}
\label{fig:factors}
\end{figure}

@@

Figure~\ref{fig:overlap} illustrates the comparison between RFNU, RFNTL 
and RFNTS. The matrices in this Figure represent the union between the 
training and the test set, where the training set is represented in 
black (negatives) and white (positives), and the test set is 
represented in green (true positives), yellow (false negatives), gray 
(true negatives) and red (false positives). This representation 
provides insights regarding where, and perhaps why, prediction errors 
occur. At this training ratio, RFNU (Figure~\ref{fig:overlap-rfnu}) 
uniformly samples the number of files per subject between 0.8$N_f$ and 
$N_f$. It provides an opportunity for ALS to be trained on the last 
files of the pipelines, while maintaining a low number of columns with 
a low training ratio. On the contrary, RFNTL 
(Figure~\ref{fig:overlap-rfntl}) samples the number of files per 
subject from 0.85$N_f$, but the probability to have a complete column 
is 0 (the one complete column in Figure~\ref{fig:overlap-rfntl} is the 
one included to prevent cold start issues), which leads to a "stripe" 
of errors at the bottom of the matrix. RFNTS 
(Figure~\ref{fig:overlap-rfnts}) doesn't have this issue, as it 
includes many complete columns in the training set, but this comes at 
the cost of several columns with a very low number of files, lower than 
60\%. For such columns, the prediction may go somewhat wrong, which explains
the reduced accuracy compared to RFNU.
